---
title: "higher order interactions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{higher order interactions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
EVAL_DEFAULT <- FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = EVAL_DEFAULT
)
```

```{r setup}
library(modsem)
```

As of yet, the `modsem` package does not support using the interaction operator `:` between two higher order constructs.
However, you can still attempt to estimate the interaction effect between two higher order constructs by specifying the interaction term in 
models using the product indicator approaches (higher order constructs are not supported at all in the `lms` and `qml` approaches yet).
The `modsem` package will then attempt to estimate the interaction effect between the two higher order constructs.

# Interaction between two higher order constructs
**WARNING:** Please note that the literature on higher order
interactions in product indicator approaches is virtually non-existant, and you
will likely need to do some experimenting. In this example you get some very 
high standard errors if you don't use the constrained approach (i.e., `method = "ca"`). 
Possibly due to the model being misspecified (**see the second example for a more realistic model**).
It might be better to use the constrained approach to estimate the interactions
but the constraints for the higher order measurement model are not yet implemented, 
and are therefore only specified for the first order measurement model. It might therefore
be a good idea to add some constraints to the model yourself (see the example below).


```{r}
tpb <- '
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN  =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3

  # Higher order constructs
  HIGH1 =~ PBC + ATT
  HIGH2 =~ INT + SN
  # Higher order interaction
  INTERACTION =~ PBC:INT + ATT:SN + PBC:SN + ATT:INT
  
  BEH =~ b1 + b2

  BEH ~ HIGH1 + HIGH2 + INTERACTION

  # Adding some constraints
  INTERACTION ~~ 0*HIGH1 + 0*HIGH2
'

est  <- modsem(tpb, data = TPB, method = "ca",
               suppress.warnings.match = TRUE)
summary(est)
```

# Interaction between a first order and a higher order construct

Here we show a more realistic example, which better fits the observed data, where we model `INT` (intention) as 
a higher order construct of `ATT` (attitude) `PBC` (perceived hehavioural control) and `SN` (subjective norm).
We then model the interaction between `INT` and `PBC` on `BEH` (behaviour). In this case we can 
see that both the constrained approach and less strict approaches yield reasonable standard errors.

```{r}
tpb <- '
  # First order constructs
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN  =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  BEH =~ b1 + b2

  # Higher order constructs
  INT =~ PBC + ATT + SN

  # Higher order interaction
  INTxPBC =~ PBC:PBC + ATT:PBC + SN:PBC
  
  # Structural model
  BEH ~ PBC + INT + INTxPBC
'

est_ca  <- modsem(tpb, data = TPB, method = "ca",
                  suppress.warnings.match = TRUE)
summary(est_ca)

est_rca  <- modsem(tpb, data = TPB, method = "rca")
summary(est_rca)

est_dblcent  <- modsem(tpb, data = TPB, method = "dblcent")
summary(est_dblcent)
```
