---
title: "interaction effects between endogenous variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{interaction effects between endogenous variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(modsem)
```

## The Problem
Interaction effects between two endogenous (i.e., dependent) variables work as you would expect for the product
indicator methods (`"dblcent", "rca", "ca", "uca"`). For the lms- and qml approach however, 
it is not as straight forward. 

The lms approach can (by default) handle interaction effects between endogenous and exogenous 
(i.e., independent) variables, 
but not interaction effects between two endogenous variables. When there is an interaction 
effect between two endogenous variables, the equations cannot easily be written in 'reduced'
form -- meaning that normal estimation procedures won't work. 

The qml approach can (by default) not handle interaction effects between two endogenous variables, 
nor interaction effects between an endogenous and an exogenous variable. The qml-approach works 
a bit differently from the lms approach (even though their fundamental equations are the same), 
so while models with interaction effects between an endogenous and an exogenous variable (in theory) can 
be estimated similarily to how they are estimated in the lms-approach, it would be a lot more computationally 
expensive than in the lms-approach.

## The Solution
This being said, there is a work-around for these limitations for both the lms- and qml-approach.
In essence, the model can be split into two parts, one linear and one non-linear. 
Basically, you can replace the covariance matrix used in the estimation of the non-linear model, 
with the model-implied covariajnce matrix from a linear model. Thus you can treat an 
endogenous variable as if it were exogenous -- given that it can be expressed in 
a linear model.

## Example 1
Let's consider the the theory of planned behaviour (TPB) where we wish to model 
the interaction effect between intention (INT) and perceived behavioural control (PBC) 
onto behaviour (BEH). The TPB model is given by the following equations: 

```{r} 
tpb <- ' 
# Outer Model (Based on Hagger et al., 2007)
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3
  BEH =~ b1 + b2

# Inner Model (Based on Steinmetz et al., 2011)
  INT ~ ATT + SN + PBC
  BEH ~ INT + PBC 
  BEH ~ PBC:INT
'
```

By default the qml-approach will not be able to estimate the model, since there 
is an interaction effect between an endogenous and and exogenous variable. 
However, we can split the model into two parts, one linear and one non-linear.
While INT is an endogenous variable, it can be expressed in a linear model, 
since it is not affected by any interaction terms: 


```{r}
tpb_linear <- 'INT ~ PBC + ATT + SN'
```

We could then remove this part from the original model, giving us: 

```{r}
tpb_nonlinear <- ' 
# Outer Model (Based on Hagger et al., 2007)
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3
  BEH =~ b1 + b2

# Inner Model (Based on Steinmetz et al., 2011)
  BEH ~ INT + PBC 
  BEH ~ PBC:INT
'
```

We could now just estimate the non-linear model, since the INT now is 
an exogenous variable. This would however not incorporate the structural model 
for INT. To address this, we can make modsem replace the covariance matrix (phi) 
of (INT, PBC, ATT, SN) with the model-implied covariance matrix from the linear model, 
whilst estimating both models simultaneously. To acheive this, we can use the 
`cov_syntax` argument in `modsem`:

```{r} 
est_qml <- modsem(tpb_nonlinear, data = TPB, cov_syntax = tpb_linear,
                  method = "qml")
summary(est_qml)
```

## Example 2
Let's now consider a similar example for the lms-approach. In the lms-approach, 
there would be no need to split the model into two parts, since the lms-approach
can handle interaction effects between endogenous and exogenous variables. 
However, the lms-approach cannot handle interaction effects between two endogenous 
variables. So we could instead consider the following model, where we 
wish to estimate the quadratic effect of INT on BEH (INT:INT):

For this example the non-linear part of the model would look the same, 
as the previous example. 


```{r}
tpb_linear <- 'INT ~ PBC + ATT + SN'
```

The non-linear part of the model would look slightly different, 
replacing `BEH ~ PBC:INT` with `BEH ~ INT:INT`:

```{r}
tpb_nonlinear <- ' 
# Outer Model (Based on Hagger et al., 2007)
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3
  BEH =~ b1 + b2

# Inner Model (Based on Steinmetz et al., 2011)
  BEH ~ INT + PBC 
  BEH ~ INT:INT
'
```

We could now just estimate the non-linear model. Now we can use the `cov_syntax` 
argument in `modsem` as in the previous example, but this time for the lms-approach:
Note: If you want the lms-approach to give as similar results as possible to
mplus, you would have to increase the number of nodes (e.g., `nodes = 100`), 
and lower the convergence criterion (e.g., `conv_crit = 1e-5`).

```{r}
est_lms <- modsem(tpb_nonlinear, data = TPB, cov_syntax = tpb_linear, method = "lms")
summary(est_lms)
```

We could also do this for the qml-approach, since it handles interaction effects 
between two exogenous variables.

```{r}
est_qml <- modsem(tpb_nonlinear, data = TPB, cov_syntax = tpb_linear, method = "qml")
summary(est_qml)
```
