---
title: "LMS and QML approaches"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LMS and QML approaches}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
EVAL_DEFAULT <- FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = EVAL_DEFAULT
)
```

```{r setup}
library(modsem)
```

# The Latent Moderated Structural Equations (LMS) and the Quasi Maximum Likelihood (QML) Approach

Both the `LMS` and `QML` approaches work on most models, but interaction effects with endogenous variables can be tricky to estimate (see the [vignette](https://modsem.org/articles/interaction_two_etas.html)). Both approaches, particularly the `LMS` approach, are computationally intensive and are partially implemented in C++ (using `Rcpp` and `RcppArmadillo`). Additionally, starting parameters are estimated using the double-centering approach, and the means of the observed variables are used to generate good starting parameters for faster convergence. If you want to monitor the progress of the estimation process, you can use `verbose = TRUE`.

## A Simple Example

Here is an example of the `LMS` approach for a simple model. By default, the `summary()` function calculates fit measures compared to a null model (i.e., the same model without an interaction term).

```{r}
library(modsem)

# ROBUST STANDARD ERRORS
m1 <- '
# Outer Model
  X =~ x1 + x2 + x3
  Z =~ z1 + z2 + z3
  Y =~ y1 + y2 + y3

# Inner Model
  Y ~ X + Z + X:Z
'

# robust.se can be used with the LMS approach, but it is recommended that 
# it is used with the QML approach instead
lms <- modsem(m1, oneInt, method = "lms", robust.se=TRUE) 
qml <- modsem(m1, oneInt, method = "qml", robust.se=TRUE)
dblcent <- modsem(m1, oneInt, method = "dblcent", estimator="MLR")


# THREE-WAY INTERACTIONS
tpb <- ' 
# Outer Model (Based on Hagger et al., 2007)
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3
  BEH =~ b1 + b2

# Inner Model (Based on Steinmetz et al., 2011)
  INT ~ ATT + SN + PBC
  BEH ~ INT + PBC 
  BEH ~ INT:PBC:SN
'

# The residual centering approach seems to work the best, but 
# that is only based on personal experimentation
rca <- modsem(tpb, TPB, method="rca")


# Standardized estimates
summary(lms1, standardized = TRUE)
```

Here is the same example using the `QML` approach:

```{r}
qml1 <- modsem(m1, oneInt, method = "qml")
summary(qml1)
```

## A More Complicated Example

Below is an example of a more complex model based on the theory of planned behavior (TPB), which includes two endogenous variables and an interaction between an endogenous and exogenous variable. When estimating more complex models with the `LMS` approach, it is recommended to increase the number of nodes used for numerical integration. By default, the number of nodes is set to 16, but this can be increased using the `nodes` argument. The `nodes` argument has no effect on the `QML` approach.

When there is an interaction effect between an endogenous and exogenous variable, it is recommended to use at least 32 nodes for the `LMS` approach. You can also obtain robust standard errors by setting `robust.se = TRUE` in the `modsem()` function.

**Note**: If you want the `LMS` approach to produce results as similar as possible to Mplus, you should increase the number of nodes (e.g., `nodes = 100`).

```{r}
# ATT = Attitude
# PBC = Perceived Behavioral Control
# INT = Intention
# SN = Subjective Norms
# BEH = Behavior
tpb <- ' 
# Outer Model (Based on Hagger et al., 2007)
  ATT =~ att1 + att2 + att3 + att4 + att5
  SN =~ sn1 + sn2
  PBC =~ pbc1 + pbc2 + pbc3
  INT =~ int1 + int2 + int3
  BEH =~ b1 + b2

# Inner Model (Based on Steinmetz et al., 2011)
  INT ~ ATT + SN + PBC
  BEH ~ INT + PBC 
  BEH ~ INT:PBC  
'

lms2 <- modsem(tpb, TPB, method = "lms", nodes = 32)
summary(lms2)

qml2 <- modsem(tpb, TPB, method = "qml")
summary(qml2, standardized = TRUE) # Standardized estimates
```
