---
title: "methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
EVAL_DEFAULT <- FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = EVAL_DEFAULT
)
```

```{r setup}
library(modsem)
```

There are a number of approaches for estimating interaction effects in SEM. In `modsem()`, the `method = "method"` argument allows you to choose which to use.

-   `"ca"` = constrained approach (Algina & Moulder, 2001)
-   `"uca"` = unconstrained approach (Marsh, 2004)
-   `"rca"` = residual centering approach (Little et al., 2006)
    -   default
-   `"dblcent"`= double centering approach (Marsh., 2013)
-   `"pind"` = basic product indicator approach (not recommended)
-   `"lms"` = The Latent Moderated Structural equations approach
    -   note: there can not be an interaction between two endogenous variables.
-   `"qml"` = The Quasi Maximum Likelihood approach.
    -   note: can only be done if you have a single endogenous (dependent) variable.
-   `"mplus"`
    -   estimates model through Mplus, if it is installed

```{r, eval = FALSE}

m1 <- '
# Outer Model
X =~ x1 + x2 + x3
Y =~ y1 + y2 + y3
Z =~ z1 + z2 + z3

# Inner model
Y ~ X + Z + X:Z 
'

modsem(m1, data = oneInt, method = "ca")
modsem(m1, data = oneInt, method = "uca")
modsem(m1, data = oneInt, method = "rca")
modsem(m1, data = oneInt, method = "dblcent")
modsem(m1, data = oneInt, method = "mplus")
modsem(m1, data = oneInt, method = "lms")
modsem(m1, data = oneInt, method = "qml")
```
